Reanalysis Step 4: Outcome Analysis with Three-Way Interaction

### 2025-08-20 reanalysis-4-outcome_analysis.qmd
df upgrade

**Three-way Cross-tabulation:**
```{r crosstab}
with(imputed_datasets[[1]], 
     ftable(dr_severity, any_treatment, ever_lapse_binary)) |> kable()
```

@@

# Pool using mice with Barnard-Rubin df
pooled_model <- pool(models)
pooled_summary <- summary(pooled_model)

# Extract results - mice::pool provides df column with Barnard-Rubin adjustment
pooled_results <- pooled_summary |>
  mutate(
    Term = term,
    Estimate = estimate,
    SE = std.error,
    t_stat = statistic,
    df = df,  # Barnard-Rubin adjusted degrees of freedom
    P_Value = p.value,
    # Recalculate CIs using t-distribution with proper df
    CI_Lower = estimate - qt(0.975, df) * std.error,
    CI_Upper = estimate + qt(0.975, df) * std.error,
    OR = exp(estimate),
    OR_CI_Lower = exp(CI_Lower),
    OR_CI_Upper = exp(CI_Upper)
  ) |>
  select(Term, Estimate, SE, t_stat, df, P_Value, CI_Lower, CI_Upper, OR, OR_CI_Lower, OR_CI_Upper)

>>>>

**Three-way Cross-tabulation:**
```{r crosstab}
# Create cross-tabulation and convert to data frame for proper display
crosstab <- with(imputed_datasets[[1]], 
                 table(dr_severity, any_treatment, ever_lapse_binary))

# Convert to data frame format
crosstab_df <- as.data.frame(crosstab)
names(crosstab_df) <- c("DR_Severity", "Treatment", "Lapse_Status", "Count")

# Pivot to wide format for better readability
library(tidyr)
crosstab_wide <- crosstab_df |>
  pivot_wider(names_from = Lapse_Status, 
              values_from = Count,
              values_fill = 0) |>
  arrange(DR_Severity, Treatment)

kable(crosstab_wide, 
      caption = "Distribution of patients by DR severity, treatment, and lapse status",
      col.names = c("DR Severity", "Treatment", "No Lapse", "Lapsed"))
```

@@

# Calculate complete-data degrees of freedom for Barnard-Rubin adjustment
# dfcom = residual degrees of freedom from complete-data model
n_obs <- nrow(models[[1]]$data)
n_params <- length(coef(models[[1]]))
dfcom <- n_obs - n_params  # This is the residual df

# Pool using mice with explicit dfcom for proper Barnard-Rubin df calculation
# mice::pool automatically applies Barnard-Rubin (1999) small-sample correction
pooled_model <- pool(models, dfcom = dfcom)
pooled_summary <- summary(pooled_model)

# Extract results - mice::pool provides df column with Barnard-Rubin adjustment
# Each parameter gets its own df based on its fraction of missing information
pooled_results <- pooled_summary |>
  mutate(
    Term = term,
    Estimate = estimate,
    SE = std.error,
    t_stat = statistic,
    df = df,  # Barnard-Rubin (1999) adjusted degrees of freedom
    P_Value = p.value,
    # Recalculate CIs using t-distribution with proper df
    CI_Lower = estimate - qt(0.975, df) * std.error,
    CI_Upper = estimate + qt(0.975, df) * std.error,
    OR = exp(estimate),
    OR_CI_Lower = exp(CI_Lower),
    OR_CI_Upper = exp(CI_Upper)
  ) |>
  select(Term, Estimate, SE, t_stat, df, P_Value, CI_Lower, CI_Upper, OR, OR_CI_Lower, OR_CI_Upper)
